Thermodynamic assessment of the Kalina Split cycle

### 2013-08-08 paper_SPLIT#4.tex
Basic addition

\author[]{Tuong-Van Nguyen\corref{cor}} \cortext[cor]{Principal corresponding author. Tel.: +45 4525 4129} \ead{tungu@mek.dtu.dk}
\author[]{Thomas Knudsen} 
\author[]{Ulrik Larsen} 
\author[]{Brian Elmegaard} 
\author[]{Fredrik Haglind}

@@

\begin{figure*}[htbp]
\centering
\includegraphics[width=0.9\linewidth]{Drawing_Kalina_Baseline.pdf}
\caption{Flowchart of the Kalina cycle process}
\label{fig:kalina_cycle}
\end{figure*}

@@

\begin{figure*}[htbp]
\centering
\includegraphics[width=1.0\linewidth]{Drawing_Kalina_Split.pdf}
\caption{Flowchart of the Kalina Split cycle process}
\label{fig:split_cycle}
\end{figure*}

@@

\pgfplotstableread[col sep=comma,header=false]{
Category 1,6,0,0
Category 2,2,2,0
Category 3,3.5,2.5,4
Category 4,3.5,2.5,4
Category 5,3.5,2.5,4
Category 6,3.5,2.5,4
Category 7,3.5,2.5,4
Category 8,3.5,2.5,4
Category 9,3.5,2.5,4
Category 10,3.5,2.5,4
Category 11,3.5,2.5,4
}\data

\pgfplotstablecreatecol[
    create col/expr={
        \thisrow{1} + \thisrow{2} + \thisrow{3} 
    }
]{sum}{\data}

\pgfplotsset{
    percentage plot/.style={
        point meta=explicit,
    every node near coord/.append style={
        align=center,
        text width=1cm
    },
        nodes near coords={
        \pgfmathtruncatemacro\iszero{\originalvalue==0}
        \ifnum\iszero=0
            \pgfmathprintnumber{\originalvalue}$\,\%$\\ \pgfmathprintnumber[fixed zerofill,precision=1]{\pgfplotspointmeta}
        \fi},
    nodes near coords align=vertical,
        yticklabel=\pgfmathprintnumber{\tick}\,$\%$,
        ymin=0,
        ymax=100,
        enlarge y limits={upper,value=0.18},
    visualization depends on={y \as \originalvalue}
    },
    percentage series/.style={
        table/y expr=\thisrow{#1}/\thisrow{sum}*100,table/meta=#1
    }
}

%%% TRY TO PLOT COMPARISON %%% 
\begin{tikzpicture}
\begin{axis}[
    axis on top,
    width=\textwidth,
    percentage plot,ybar,bar width=0.1cm,
    enlarge x limits=0.25,
    symbolic x coords={Category 1,Category 2,Category 3,Category 4,Category 5,Category 6,Category 7,Category 8,Category 9,Category 10,Category 11},
    xtick=data
]
\addplot table [percentage series=1] {\data};
\addplot table [percentage series=2] {\data};
\addplot table [percentage series=3] {\data};
%\addplot table [percentage series=4] {\data};
\end{axis}
\end{tikzpicture}

@@

\subsection{Energy analysis}

\subsection{Exergy analysis}

>>>>

%\author[]{Tuong-Van Nguyen\corref{cor}} \cortext[cor]{Principal corresponding author. Tel.: +45 4525 4129} \ead{tungu@mek.dtu.dk}
%\author[]{Thomas Knudsen} 
%\author[]{Ulrik Larsen} 
%\author[]{Brian Elmegaard} 
%\author[]{Fredrik Haglind}

@@

\begin{figure*}[htbp]
\centering
\includegraphics[scale=0.65]{Drawing_Kalina_Baseline.pdf}
\caption{Flowchart of the Kalina cycle process}
\label{fig:kalina_cycle}
\end{figure*}

@@

\begin{figure*}[htbp]
\centering
\includegraphics[scale=0.65]{Drawing_Kalina_Split.pdf}
\caption{Flowchart of the Kalina Split cycle process}
\label{fig:split_cycle}
\end{figure*}

@@

@@

\subsection{Comparison of the 4 configurations}

%%% TRY TO PLOT COMPARISON %%%

\begin{tikzpicture}
  \centering
  \begin{axis}[
    xbar,
    bar width=0.2cm,
    legend style={at={(0.5,-0.15)},
    %enlarge x limits={upper,value=0.1},
    xmin=0,xmax=120,
    anchor=north,legend columns=-1},
    symbolic y coords={Cooling water,Shaft,Heat recovery system,Mixers,Recuperators,Condensers,Separator,Throttling valve,Pumps,Turbines},
    ytick=data,
	%enlarge y limits={upper,value=0.1},
    nodes near coords,
    nodes near coords align=horizontal,
	    width=0.7\textwidth,
		height=0.7\textheight 
    ]
    \addplot coordinates {(75.4064,Cooling water) (89.7961,Shaft) (94.4597,Heat recovery system)
      (96.6786,Mixers) (97.5600,Recuperators) (98.2339,Condensers)
      (98.6138,Separator) (98.9129,Throttling valve) (99.0970,Pumps)(99.0970,Turbines)};
	      \addplot coordinates {(75.4064,Cooling water) (89.7961,Shaft) (94.4597,Heat recovery system)
      (96.6786,Mixers) (97.5600,Recuperators) (98.2339,Condensers)
      (98.6138,Separator) (98.9129,Throttling valve) (99.0970,Pumps)(99.0970,Turbines)};
    \addplot coordinates {(30.5101,Cooling water) (34.5384,Shaft) (36.3324,Heat recovery system)
      (37.3570,Mixers) (37.9158,Recuperators) (38.3514,Condensers)
      (38.6484,Separator) (38.9125,Throttling valve)(99.0970,Pumps) (39.1067,Turbines)};
	      \addplot coordinates {(75.4064,Cooling water) (89.7961,Shaft) (94.4597,Heat recovery system)
      (96.6786,Mixers) (97.5600,Recuperators) (98.2339,Condensers)
      (98.6138,Separator) (98.9129,Throttling valve) (99.0970,Pumps)(99.0970,Turbines)};
    \legend{Kalina cycle,Kalina cycle with reheat, Kalina Split cycle, Kalina Split cycle with reheat}
  \end{axis}
\end{tikzpicture}

\subsection{Energy and Exergy flow diagrams}

\subsection{Effect of the boundary conditions}

### 2013-08-08 paper_SPLIT#5.tex
Basic addition

\author[]{Tuong-Van Nguyen\corref{cor}} \cortext[cor]{Principal corresponding author. Tel.: +45 4525 4129} \ead{tungu@mek.dtu.dk}
\author[]{Thomas Knudsen} 
\author[]{Ulrik Larsen} 
\author[]{Brian Elmegaard} 
\author[]{Fredrik Haglind}

@@

	\begin{equation}
	 \dot{Q}_{heat}=\dot{W}+\dot{Q}_{cw}
	\end{equation}

	where $\dot{Q}_{heat}$ is the energy recovered as heat from the waste heat source, $\dot{W}$ the electric power delivered by the plant, and $\dot{Q}_{cool}$ the energy losses due to heat rejection to the environment with the cooling water.

@@

	\begin{equation}
		\dot{E}^{Q}_{heat}=\dot{E}^{W}+\dot{E}^{Q}_{cw}+\dot{E}_{d}
	\end{equation}

	where $\dot{E}^{Q}_{heat}$ represents the exergy decrease of the heat source, which is used as fuel to the power cycle system. $\dot{E}^{W}$ denotes the exergy associated with the net power produced, $\dot{E}^{Q}_{cool}$ the exergy discharged to the environment with the exergy gain of the cooling water, and $\dot{E}_{d}$ the exergy destroyed in the plant. The dead state is taken to the environmental conditions and the reference environment of Szargut \cite{Szargut1998} is considered for the chemical exergy calculations.

@@

\subsection{Energy analysis}

\subsection{Exergy analysis}

>>>>

\author[]{Tuong-Van Nguyen\corref{cor}} \cortext[cor]{Principal corresponding author. Tel.: +45 4525 4129} \ead{tungu@mek.dtu.dk}
\author[]{Thomas Knudsen} 
\author[]{Ulrik Larsen} 
\author[]{Fredrik Haglind} 
\author[]{Brian Elmegaard}

@@

	\begin{equation}
	 \dot{Q}_{heat}=\dot{W}+\dot{Q}_{cw}+\dot{Q}_{exh}
	\end{equation}

	where $\dot{Q}_{heat}$ is the heat input to the system with the waste heat source, $\dot{W}$ the electric power delivered by the plant, $\dot{Q}_{cool}$ and $\dot{Q}_{exh}$ the energy losses due to heat rejection to the environment with the cooling water and exhaust gases.

@@

	\begin{equation}
		\dot{E}^{Q}_{heat}=\dot{E}^{W}+\dot{E}^{Q}_{cw}+\dot{E}^{Q}_{exh}+\dot{E}_{d}
	\end{equation}

	where $\dot{E}^{Q}_{heat}$ represents the exergy input to the power cycle with the heat source. $\dot{E}^{W}$ denotes the exergy associated with the net power produced, $\dot{E}^{Q}_{cool}$ and $\dot{E}^{Q}_{exh}$ the exergy discharged to the environment with the cooling water and exhaust gases, and $\dot{E}_{d}$ the exergy destroyed in the plant. The dead state is taken to the environmental conditions and the reference environment of Szargut \cite{Szargut1998} is considered for the chemical exergy calculations.

@@

\subsection{Comparison of the 4 configurations}

The exergy destruction in the Kalina cycle amount to 1726 (\emph{Case 1}), 1681 (\emph{Case 2}), 1662 (\emph{Case 3}) and 1575 (\emph{Case 4}) kW. The exergy analysis indicates that the implementation of the reheat and split cycle configurations result in smaller exergy destruction and losses by 2.5--4.9\% and 2.7--5.1\% (Figure \ref{fig:comparison_exergy_destruction_kalina}). In any case, the greatest irreversibilities take place (1) in the turbine(s), (2) in the heat recovery system, (3) in the condensers and recuperators. The first ones are related to the inefficiencies of the expansion process, while the others are caused by heat transfer across the temperature differences between the working fluid and the exhaust gases/cooling water. The Kalina Split cycle is characterised by a significant reduction of the exergy destruction in heat transfer processes, mainly because of the better temperature match between the heat source and cold reservoir in the intermediate pressure condenser (36--38\%) and heat recovery system (4.8--24\%). In contrast, the exergy destruction in the recuperators is higher, because more heat is transferred in these heat exchangers. The exergy losses associated with the rejection of the exhaust gases to the environment are constant and equal to \textcolor{blue}{(v\ae rdi)}, since the comparison of the four Kalina cycle configurations is based on fixed temperature and pressure of the exhaust gases at the outlet of the heat recovery system. On the opposite, the exergy losses with cooling water are slightly higher.

\begin{figure*}[htpb]
\centering
\begin{tikzpicture}
  \centering
  \begin{axis}[
    xbar,
    bar width=0.1cm,%xlabel={\#percentage} %nodes near coords={(\coordindex \#percentage)}, nodes near coords align=horizontal,
	legend style={at={(0.5,-0.15)},     %enlarge x limits={upper,value=0.1},
    xmin=0,xmax=600,
	anchor=north,legend columns=-1},
	symbolic y coords={Turbines,Pumps,Throttling valve,Condensers,Recuperators,Mixers,Heat recovery,Shaft,Cooling water},     
	x tick label style={/pgf/number format/1000 sep=},
	ytick=data, %	enlarge y limits={upper,value=0.1}, 
	    width=0.5\textwidth,
		height=0.5\textheight 
    ]
	\addplot coordinates {(497.0003749,Turbines) (37.31435888,Pumps) (0.694136421,Throttling valve) (405.6245141,Condensers) (238.9098024,Recuperators) (31.19770433,Mixers) (426.0626603,Heat recovery) (88.76973684,Shaft) (124.1733785,Cooling water)};
	\addplot coordinates {(458.8176706,Turbines) (29.01391041,Pumps) (1.613922611,Throttling valve) (396.1664419,Condensers) (253.3386248,Recuperators) (17.99282833,Mixers) (435.6833497,Heat recovery) (88.31578947,Shaft) (121.90394,Cooling water)};	
	\addplot coordinates {(528.364695,Turbines) (30.40378801,Pumps) (0.971456159,Throttling valve) (342.5550964,Condensers) (261.5016632,Recuperators) (4.236863901,Mixers) (405.6368272,Heat recovery) (88.10682743,Shaft) (137.3728447,Cooling water)};	
	\addplot coordinates {(513.9937499,Turbines) (28.51276797,Pumps) (1.684363437,Throttling valve) (330.775187,Condensers) (272.0408547,Recuperators) (3.999427225,Mixers) (330.8117831,Heat recovery) (92.93083621,Shaft) (135.7513645,Cooling water)};	  
    \legend{Kalina cycle,Kalina cycle with reheat, Kalina Split cycle, Kalina Split cycle with reheat}
 \end{axis}
\end{tikzpicture}
\caption{Comparison of the exergy destruction and losses in the four Kalina cycle configurations (Kalina cycle (Case 1), Kalina cycle with reheat (Case 2), Kalina Split cycle (Case 3), Kalina Split cycle with reheat (Case 4)). The exergy losses associated with the rejection of the exhaust gases are not shown, since they are equal in the four cases.}
\label{fig:comparison_exergy_destruction_kalina}
\end{figure*}

\subsection{Energy and Exergy flow diagrams}

\subsection{Effect of the boundary conditions}

\subsection{Entropy generation in the heat recovery system}
